---
title: "Assignment 2"
author: "Tareq Al Hossain Tomal"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, message = FALSE)
```

#install.packages("readr")  
#install.packages("ggplot2") 
#install.packages("knitr")
#install.packages("tidyverse")
#install.packages("naniar")
#install.packages("writexl")



# Packages

Loading required packages:

```{r message=FALSE, warning=FALSE, echo=FALSE}
library(knitr)
library(ggplot2)
library(dplyr)
library(readr)
library(tidyverse)
library(naniar)
library(writexl)
```

------------------------------------------------------------------------

# Import the Data

```{r}
data <- read.csv("Kaggle - Pima Indians Diabetes Database.csv")
head(data)

```

Here is what the columns in the data mean:

| Column | Details |
|--|----|
| Pregnancies | Number of times pregnant |
| Glucose | Plasma glucose concentration a 2 hours in an oral glucose tolerance test |
| BloodPressure | Diastolic blood pressure (mm Hg) |
| SkinThickness | Triceps skin fold thickness (mm) |
| Insulin | 2-Hour serum insulin (mu U/ml) |
| BMI | Body mass index (weight in kg/(height in m)\^2) |
| DiabetesPedigreeFunction | Diabetes pedigree function |
| Age | Age (years) |
| Outcome | Class variable (0=No or 1=Yes) |

------------------------------------------------------------------------

## Question 1 (10 marks)

**`Outcome` is a categorical variable, 1 means Yes, 0 means No. So, convert it into a factor data type. Now, calculate mean, standard deviation, minimum, 1st quartile, median, 3rd quartile, maximum, and inter-quartile range of all the numeric columns in the dataset. Do you see any anomalies? Write your comment.**

```{r}

colnames(data) <- make.names(colnames(data), unique = TRUE)


data$Outcome <- as.factor(data$Outcome)

numeric_summary <- data %>%
  select(-Outcome) %>%  # Exclude Outcome column
  summarise_all(list(
    mean = ~mean(. , na.rm = TRUE),
    sd = ~sd(. , na.rm = TRUE),
    min = ~min(. , na.rm = TRUE),
    q1 = ~quantile(. , 0.25, na.rm = TRUE),
    median = ~median(. , na.rm = TRUE),
    q3 = ~quantile(. , 0.75, na.rm = TRUE),
    max = ~max(. , na.rm = TRUE),
    iqr = ~IQR(. , na.rm = TRUE)
  ))

# View the summary statistics
numeric_summary


```



**How many people have Outcome = Yes?**

```{r}
outcome_yes_count <- sum(data$Outcome == 1)


outcome_yes_count


```


## Question 2 (40 marks)

**Create histograms for all the numeric variables in the data using ggplot2.**


```{r}

data_long <- data %>%
  select(-Outcome) %>% 
  gather(key = "Variable", value = "Value")


ggplot(data_long, aes(x = Value)) +
  geom_histogram(binwidth = 10, fill = "skyblue", color = "black", alpha = 0.7) +
  facet_wrap(~ Variable, scales = "free", ncol = 3) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables", x = "Value", y = "Frequency")



```

## Question 3 (20 marks)

**Create boxplot of Glucose, Insulin, BMI, and Age by Outcome using ggplot2.**

```{r}
ggplot(data, aes(x = Outcome, y = Glucose, fill = Outcome)) +
  geom_boxplot() +
  labs(title = "Boxplot of Glucose by Outcome", x = "Outcome", y = "Glucose") +
  theme_minimal()

ggplot(data, aes(x = Outcome, y = Insulin, fill = Outcome)) +
  geom_boxplot() +
  labs(title = "Boxplot of Insulin by Outcome", x = "Outcome", y = "Insulin") +
  theme_minimal()

ggplot(data, aes(x = Outcome, y = BMI, fill = Outcome)) +
  geom_boxplot() +
  labs(title = "Boxplot of BMI by Outcome", x = "Outcome", y = "BMI") +
  theme_minimal()

ggplot(data, aes(x = Outcome, y = Age, fill = Outcome)) +
  geom_boxplot() +
  labs(title = "Boxplot of Age by Outcome", x = "Outcome", y = "Age") +
  theme_minimal()


```


## Question 4 (10 marks)

**Replace 0 with NA in the variables where a value of 0 does not make sense.**

```{r}
data$Glucose[data$Glucose == 0] <- NA
data$BloodPressure[data$BloodPressure == 0] <- NA
data$SkinThickness[data$SkinThickness == 0] <- NA
data$Insulin[data$Insulin == 0] <- NA
data$BMI[data$BMI == 0] <- NA


summary(data)


```


## Question 5 (10 marks)

**Use naniar package to inspect number of missing values in the data after replacing 0s with NAs**

```{r}
gg_miss_var(data)  

miss_summary <- miss_var_summary(data)

miss_summary


```


## Question 6 (10 marks)

**Export the final cleaned data in an Excel file for later use.**

```{r}
write_xlsx(data, "cleaned_pima_diabetes_data.xlsx")

```


